### 2024-03-12 output/paper.qmd
abstract, started tests

---
title: "Predicting Political Support in the United States"
# subtitle: "My subtitle if needed"
author: 
  - Janssen Myer Rambaud
  - Timothius Prajogi
thanks: "Code and data are available at: https://github.com/prajogt/predicting_political_support_us.git ."
date: today
date-format: long
abstract: "Using data gathered from the 2022 Cooperative Election Study (CES), we navigate through the various possible factors that influence a US citizen's voting decision. By examining a representative sample of 60 000 Americans, we can discern patterns in beliefs and personal backgrounds—ranging from gun control and student loan stances to race, religion, and educational background. This paper discusses the impact these individual factors may have towards predicting a voter's alignment, as well as how multiple factors can coalign to portray a more substantial impact. We aim to highlight these correlations and their potential implications with future elections."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction
include Estimand in here

# Data
needs measurement

# Model

# Results

# Discussion

# Appendix

### 2024-03-12 scripts/00_simulate_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
-# Purpose: Simulates data for survey results regarding financial attitudes
+# Purpose: Simulates data for predicting US political support based on demographic and socio-economic factors
 # Author: Janssen Myer Rambaud & Timothius Prajogi
 # Data: March 11 2024
-# Contact: janssen.rambaud@mail.utoronto.ca timothius.prajogi@mail.utoronto.ca
+# Contact: janssen.rambaud@mail.utoronto.ca, timothius.prajogi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
 # Assumptions:
@@ -19,7 +19,7 @@ num_obs <- 1000
 # this is % of electorate to calculate the eligible voters (RIGHT SIDE OF COLUMN CES2022 PG 24-25)
 us_political_preferences <- tibble(
   registered_to_vote = sample(c("Yes", "No", "Don't know"), size = num_obs, replace = TRUE, prob = c(0.932, 0.062, 0.006)),
-  birth_year = sample(1940:2006, size = num_obs, replace = TRUE),
+  birth_year = sample(1940:2004, size = num_obs, replace = TRUE),
   gender = sample(c(0, 1), size = num_obs, replace = TRUE, prob = c(0.48, 0.52)),
   education = sample(c(0, 1, 2, 3), size = num_obs, replace = TRUE, prob = c(0.29, 0.29, 0.26, 0.16)),
   race = sample(c(0, 1, 2, 3, 4), size = num_obs, replace = TRUE, prob = c(0.75, 0.09, 0.09, 0.04, 0.02)),
@@ -98,3 +98,33 @@ us_political_preferences <- tibble(
   ) |>
   select(registered_to_vote, birth_year, age_group, gender, education, race, religion, gun_ownership, education_loans, supports_biden)
 
+
+#### Testing the data ####
+
+# Test registered_to_vote
+table(us_political_preferences$registered_to_vote %in% c("Yes", "No", "Don't know"))
+
+# Test birth_year
+min(us_political_preferences$birth_year) == 1940
+max(us_political_preferences$birth_year) == 2004
+
+# Test gender
+table(us_political_preferences$gender %in% c("Male", "Female"))
+
+# Test education
+table(us_political_preferences$education %in% c("High school or less", "Some college", "College", "Post-grad"))
+
+# Test race
+table(us_political_preferences$race %in% c("White", "Black", "Hispanic", "Asian", "Other"))
+
+# Test religion
[…]

### 2024-03-12 output/paper.qmd
Start Paper

@@
 #| message: false
+#| echo: false
 
+#### Workspace setup ####
+library(knitr)
 library(tidyverse)
+
 ```
@@
 
-# Results
+```{r}
 
-# Discussion
+cm <- c('raceWhite' = 'White',
+        'raceNative American' = 'Native American', 
+        'raceMultiracial' = 'Multiracial',
+        'raceMiddle Eastern' = 'Middle Eastern',
+        'raceHispanic' = 'Hispanic',
+        'raceBlack' = 'Black',
+        'genderMale' = 'Male')
 
-# Appendix
+modelsummary(
+  list(
+    "Support Biden" = political_preferences_gender_race
+  ),
+  statistic = "mad"
+  )
+
+modelplot(political_preferences_gender_race, conf_level = 0.9, coef_map = cm) + labs(x = "90 per cent credibility interval")
+```
+
+## Results
+
+## Discussion
+
+## Conclusion
+
+## Appendix
+
+## References

### 2024-03-12 scripts/03_model.R

@@ -26,7 +26,7 @@ ces2022_1_reduced <-
   ces2022_1 |>
   slice_sample(n = 1000) |>
   mutate(
-    voted_for = as_factor(voted_for)
+    voted_for = if_else(voted_for == "Biden", 1, 0)
   )
 
 ces2022_2_reduced <-
@@ -35,23 +35,12 @@ ces2022_2_reduced <-
   mutate(
     voted_for = as_factor(voted_for)
   )
-
-ces2022_gender_reduced <-
-  ces2022_1 |>
-  filter(
-    gender != "Other",
-    race != "Other"
-  ) |>
-  mutate(
-    voted_for = if_else(voted_for == "Biden", 0, 1)
-  ) |>
-  slice_sample(n = 1000)
   
 # Model
 political_preferences_gender_race <-
   stan_glm(
     voted_for ~ gender + race,
-    data = ces2022_gender_reduced,
+    data = ces2022_1_reduced,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = 

### 2024-03-12 scripts/03_model.R
update gender education results
save to images, also swapped the graph colours to make it accurate to the actual candidates

@@ -64,5 +64,8 @@ saveRDS(
 #    fill = "Voted for"
 #  ) +
 #  coord_flip() +
-#  scale_fill_brewer(palette = "Set1") +
+#  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
 #  theme(legend.position = "bottom")
+
+#ggsave('output/images/gender_education_results.jpg')
+

### 2024-03-12 output/paper.qmd
incl. library model summary
Add cultural, race, and importance models

--- 2024-03-12 output/paper.qmd
@@
 title: "Predicting Political Support in the United States"
-# subtitle: "My subtitle if needed"
+subtitle: "The Effects of Culture and Key Political Choices"
 author: 
@@
 #| message: false
+#| echo: false
 
+#### Workspace setup ####
+library(knitr)
 library(tidyverse)
+library(modelsummary)
 ```
@@
 
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+#| echo: false
+
+# Load in data
+ces2022_1 <- read_csv("data/cleaned_ces2022_1.csv")
+ces2022_2 <- read_csv("data/cleaned_ces2022_2.csv")
+```
+
+```{r}
+ces2022_1 |>
+  ggplot(aes(x = education, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender)) +
+  theme_minimal() +
+  labs(
+    x = "Highest education",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+
+```
+
 # Model
 
-# Results
+```{r}
+# Load in models
+political_preferences_cultural <-
+  readRDS(file = "models/political_preferences_cultural.rds")
 
-# Discussion
+political_preferences_race <-
+  readRDS(file = "models/political_preferences_race.rds")
 
-# Appendix
+political_preferences_importances <-
+  readRDS(file = "models/political_preferences_importances.rds")
+```
+
+```{r}
+#| eval: true
+#| fig-cap: "Logistic regression probability results with Gender, Age, Race, and Religion as Predictors"
+#| include: true
+#| label: fig-modelcultural
+#| message: false
+#| warning: false
+
+modelsummary(
+  list(
+    "Support Biden" = political_preferences_cultural
+  ),
+  statistic = "mad"
+  )
+
+```
+
+```{r}
+modelsummary(
+  list(
+    "Support Biden" = political_preferences_importances
+  ),
+  statistic = "mad"
+  )
+
+```
+
+## Results
+
+## Discussion
+
+Although there seems to be a strong correlation between if a voter
+identifies as an other gender to supporting Biden, and a strong correlation
+between if a voter identifies as an other race to supporting Biden,
+but these must be considered carefully as within the sample (with seed 302), 
+only 5 and 1 voters identified in those categories respectively. To get a more
+accurate reading in regards to these features, it would be necessary to train
+a model which includes much more of these demographics.
+
+## Appendix
+
+```{r}
+cm <- c('genderMale' = 'Male',
+        'raceWhite' = 'White',
+        'raceNative American' = 'Native American', 
+        'raceMultiracial' = 'Multiracial',
+        'raceMiddle Eastern' = 'Middle Eastern',
+        'raceHispanic' = 'Hispanic',
+        'raceBlack' = 'Black',
+        'raceAsian' = 'Asian',
+        'religionNone' = 'Non-Religious',
+        'religionProtestant' = 'Protestant',
+        'religionAgnostic' = 'Agnostic',
+        'religionAtheist' = 'Atheist',
+        'religionRoman Catholic' = 'Roman Catholic',
+        'religionJewish' = 'Jewish',
+        'religionOther' = 'Other'
+        )
+
+modelplot(political_preferences_gender_race_religion, conf_level = 0.9, coef_map = cm) +
+  labs(x = "90 per cent credibility interval")
+```
+
+```{r}
+modelplot(political_preferences_gender_race_religion, 
+          conf_level = 0.9, 
+          coef_map = c("raceOther" = "Other Race", "genderOther" = "Other Gender (Does not Identify as Male or Female)")) + 
+  labs(x = "90 per cent credibility interval")
+```
+
+## References

### 2024-03-12 scripts/03_model.R

@@ -27,19 +27,21 @@ ces2022_1_reduced <-
   slice_sample(n = 1000) |>
   mutate(
     voted_for = if_else(voted_for == "Biden", 1, 0)
+    # voted_for = as_factor(voted_for)
   )
 
 ces2022_2_reduced <-
   ces2022_2 |>
   slice_sample(n = 1000) |>
   mutate(
-    voted_for = as_factor(voted_for)
+    voted_for = if_else(voted_for == "Biden", 1, 0)
+    # voted_for = as_factor(voted_for)
   )
   
-# Model
-political_preferences_gender_race <-
+# Model with gender, age_group, race, religion, education
+political_preferences_cultural <-
   stan_glm(
-    voted_for ~ gender + race,
+    voted_for ~ gender + age_group + race + religion,
     data = ces2022_1_reduced,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -49,8 +51,8 @@ political_preferences_gender_race <-
   )
 
 saveRDS(
-  political_preferences_gender_race,
-  file = "output/models/political_preferences_gender_race.rds"
+  political_preferences_cultural,
+  file = "output/models/political_preferences_cultural.rds"
 )
 
 #ces2022_1 |>
@@ -69,3 +71,36 @@ saveRDS(
 
 #ggsave('output/images/gender_education_results.jpg')
 
+# Model with only race
+political_preferences_race <-
+  stan_glm(
+    voted_for ~ race,
+    data = ces2022_1_reduced,
+    family = binomial(link = "logit"),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = 
+      normal(location = 0, scale = 2.5, autoscale = TRUE),
+    seed = 302
+  )
+
+saveRDS(
+  political_preferences_race,
+  file = "output/models/political_preferences_race.rds"
+)
+
+# Model with education, gunown, edloan (political differences)
+political_preferences_importances <-
+  stan_glm(
+    voted_for ~ education + household_gun_ownership + student_loan_status,
+    data = ces2022_2_reduced,
+    family = binomial(link = "logit"),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = 
+      normal(location = 0, scale = 2.5, autoscale = TRUE),
+    seed = 302
+  )
+
+saveRDS(
+  political_preferences_importances,
+  file = "output/models/political_preferences_importances.rds"
+)

### 2024-03-12 scripts/03_model.R
Clean up model script

@@ -4,7 +4,7 @@
 # Date: 23 January 2024
 # Contact: tim.prajogi@mail.utoronto.ca
 # License: MIT
-# Prerequisites: 01_download_data
+# Prerequisites: 01_download_data, 02_clean_data
 
 #### Workplace setup ####
 
@@ -54,22 +54,6 @@ saveRDS(
   political_preferences_cultural,
   file = "output/models/political_preferences_cultural.rds"
 )
-
-#ces2022_1 |>
-#  ggplot(aes(x = education, fill = voted_for)) +
-#  stat_count(position = "dodge") +
-#  facet_wrap(facets = vars(gender)) +
-#  theme_minimal() +
-#  labs(
-#    x = "Highest education",
-#    y = "Number of respondents",
-#    fill = "Voted for"
-#  ) +
-#  coord_flip() +
-#  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
-#  theme(legend.position = "bottom")
-
-#ggsave('output/images/gender_education_results.jpg')
 
 # Model with only race
 political_preferences_race <-

### 2024-03-12 output/paper.qmd
Update references

@@
 # Data
+
+These delay statistics were downloaded, cleaned, parsed, analyzed, and visualized using R [@citeR], a statistical programming language, with package support from `tidyverse` [@citetidyverse], a collection of libraries which included the following packages that were utilized:
+
+* `ggplot2` [@citeggplot] 
+* `dplyr` [@citedplyr] 
+* `readr` [@citereadr] 
+* `tibble` [@citetibble]
+
+The data was retrieved from Harvard's CCES Dataverse database, using the `dataverse` [@citedataverse] package.
+
+In this report we consider data collected from the 2022 CES survey, stored by Dataverse as ***Cooperative Election Study Common Content*** [@citecesdata].
+
 needs measurement
@@
 
-modelplot(political_preferences_gender_race_religion, conf_level = 0.9, coef_map = cm) +
+modelplot(political_preferences_cultural, conf_level = 0.9, coef_map = cm) +
   labs(x = "90 per cent credibility interval")
@@
 ```{r}
-modelplot(political_preferences_gender_race_religion, 
+modelplot(political_preferences_cultural, 
           conf_level = 0.9, 

### 2024-03-12 output/paper.qmd
Add model section

@@
 bibliography: references.bib
+editor: visual
 ---
@@
 library(modelsummary)
+library(rstanarm)
+
 ```
@@
 
-In this report we consider data collected from the 2022 CES survey, stored by Dataverse as ***Cooperative Election Study Common Content*** [@citecesdata].
+In this report we consider data collected from the 2022 CES survey, stored by Dataverse as *Cooperative Election Study Common Content* [@citecesdata].
 
-needs measurement
+needs measurement here
 
@@
 
+To create the generalized linear model we made use of the `rstanarm` [@citerstanarm] package.
+
+In particular the models we had created are logistic regression, all prediciting who the participant voted for during the 2020 US Presidential Election.
+
 ```{r}
@@
 
+The first model that was created was intended to discover whether we can forecast who a respondent was likely to vote for, knowing their cultural background, including their age, gender, race, and religion.
+
+We expect that due to various differences in experiences due to this background, whether it be because of a generational difference, a difference in culture between countries, or a difference in beliefs in religions, a voter who identifies in those demographics would vote differently, voting for the candidate who most undestands their experiences.
+
+The second model only forecasts who the respondent was likely to vote for based on race, to extract how much of the intercept in the previous model can be explained by a different in race.
+
+Finally, the last model aims to forecast a respondents vote knowing their prominent political ideals including education, gun ownership status, and student loan status.
+These are common political standpoints in the U.S..
+Gun policy and student loan forgiveness is often a key political issue that separates the candidates.
+For the 2020 election, the candidates we are predicting for, Joe Biden and Donald Trump had opposing positions on student loan forgiveness.
+Whilst Trump was opposed to student loan forgiveness, backing incone-based repayment plans instead [@citetrumploanstance], Biden proposed to forgive a large amount of undergraduate tuition-related student debt [@citeballotpedia].
+They also opposed each other on gun laws, with Trump advocating for the right to bear arms and Biden advocating gun safety policies aimed at holding manufacturers accountable [@citeballotpedia].
+
+As such, we expect that there would be a strong correlation with the respondents gun ownership and student loan status to the candidate they supported in the 2020 election.
+
+TODO add to refs
+
+https://www.washingtonpost.com/politics/interactive/2023/presidential-candidates-2024-policies-issues/donald-trump-student-loans-education/
+
+https://ballotpedia.org/2020_presidential_candidates_on_student_loan_debt
+https://www.rand.org/research/gun-policy/key-findings/gun-policy-in-america.html
+
 ```{r}
-#| eval: true
-#| fig-cap: "Logistic regression probability results with Gender, Age, Race, and Religion as Predictors"
-#| include: true
-#| label: fig-modelcultural
+#| label: tbl-modelsummarycultural
+#| tbl-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their gender, age, race, and religion"
 #| message: false
 #| warning: false
+#| echo: false
 
@@
 
+\newpage
 ## References

### 2024-03-12 output/paper.qmd
introduction complete

@@
 include Estimand in here
+
+<!-- The introduction is self-contained and tells a reader everything they need to know including:
+
+1) broader context to motivate;
+
+2) some detail about what the paper is about;
+
+3) a clear gap that needs to be filled;
+
+4) what was done;
+
+5) what was found;
+
+6) why it is important;
+
+7) the structure of the paper.
+
+A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total. -->
+
+Voting for the leader of your country is one of the biggest decisions an American citizen can make.
+It decides many of the decisions that will be made throughout that four year term as well as how their life will change in that period of time.
+There may be changes that this elected government makes that they believe to be a great decision, but it may also be seen as terrible one in the eyes of others.
+This is why it is important to understand there are many factors that influence a voter's decision.
+They vote for the candidate who they believe will make the best decisions for them and their country.
+In this paper, we hope to show the impact of various factors on a voter's decision, and how these factors can be used to predict a voter's alignment.
+From personal stances and religious beliefs to political stances, or to biological factors, we hope to show how these factors can be used to predict a voter's alignment.
+
+The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump (Gambino 2020).
+Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors.
+We will be using a representative sample of 60,000 Americans to discern patterns in these voters' responses  and how their personal backgrounds may have influenced their voting decision.
+We will be examining the impact of various factors such as: gender, race, educational background, religon, and personal stances on key political issues such as gun control and student loans.
+Our main focus is to compare the correlations in the 2022 CES survey data to the 2020 CES survey data, and to see if there are any significant correlations that can be used to predict a voter's alignment.
+Using these understandings, we hope to highlight the potential implications these correlations may have for this upcoming 2024 election.
+
+With this context and information in mind, we developed this paper to be structured as follows.
+We will first present the data and the models we used to analyze the data.
+We will then present the results of our analysis, and discuss the implications of these results.
+Finally, we will conclude with a summary of our findings and the potential implications of these findings for the upcoming 2024 election.
+To go in more detail, after downloading and cleaning the 2022 CES survey data, we used the `rstanarm` [@citerstanarm] package to create a generalized linear model to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors, as listed above.
+We then used the `modelsummary` [@citemodelsummary] package to summarize the results of our model.
+We found that there were strong correlations, with a strong example of this being the voter's personal stances on student loan forgiveness and gun ownership.
+With the two main parties, Democrats and Republicans, having opposing stances on these issues, we found that there was a strong correlation between these stances and who a voter supported.
+As Biden was in favor of student loan forgiveness, but opposed to gun ownership, while Trump was opposed to student loan forgiveness, but in favor of gun ownership, we discovered a correlation between a voter's ties to these stances because they directly oppose each other.
+This is just one example of the many correlations we found, and we will discuss these in more detail in the results section.
+These findings are important because it shows how a leading presidential party's stance on a certain topic has the potential to sway a voter's decision in the event that they align with that stance.
+This is important because it shows how a leading presidential party's stance on a certain topic has the chance to adjust and adapt to certain factors to possibly sway some more votes in their favour.
+We will also discuss the implications of these findings, and how they may be used to predict a voter's alignment in the upcoming 2024 election.
+
+TODO: cite https://www.theguardian.com/us-news/2020/nov/07/joe-biden-wins-us-election-donald-trump-loses-final-result-2020
 
@@
   coord_flip() +
-  scale_fill_brewer(palette = "Set1") +
+  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
   theme(legend.position = "bottom")
@@
 
-In particular the models we had created are logistic regression, all prediciting who the participant voted for during the 2020 US Presidential Election.
+In particular the models we had created are logistic regression, all predicting who the participant voted for during the 2020 US Presidential Election.
 
@@
 For the 2020 election, the candidates we are predicting for, Joe Biden and Donald Trump had opposing positions on student loan forgiveness.
-Whilst Trump was opposed to student loan forgiveness, backing incone-based repayment plans instead [@citetrumploanstance], Biden proposed to forgive a large amount of undergraduate tuition-related student debt [@citeballotpedia].
-They also opposed each other on gun laws, with Trump advocating for the right to bear arms and Biden advocating gun safety policies aimed at holding manufacturers accountable [@citeballotpedia].
+Whilst Trump was opposed to student loan forgiveness, backing incone-based repayment plans instead ()[@citetrumploanstance], Biden proposed to forgive a large amount of undergraduate tuition-related student debt ()[@citeballotpedia].
+They also opposed each other on gun laws, with Trump advocating for the right to bear arms and Biden advocating gun safety policies aimed at holding manufacturers accountable ()[@citeballotpedia].
 

### 2024-03-14 output/paper.qmd
Added to data, fixed 02_clean_data.R because the age 64 was a typo.

@@
 # Introduction
-include Estimand in here
-
-<!-- The introduction is self-contained and tells a reader everything they need to know including:
-
-1) broader context to motivate;
-
-2) some detail about what the paper is about;
-
-3) a clear gap that needs to be filled;
-
-4) what was done;
-
-5) what was found;
-
-6) why it is important;
-
-7) the structure of the paper.
-
-A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total. -->
 
@@
 Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors.
-We will be using a representative sample of 60,000 Americans to discern patterns in these voters' responses  and how their personal backgrounds may have influenced their voting decision.
+We will be using a representative sample of 60000 Americans to discern patterns in these voters' responses  and how their personal backgrounds may have influenced their voting decision.
 We will be examining the impact of various factors such as: gender, race, educational background, religon, and personal stances on key political issues such as gun control and student loans.
-Our main focus is to compare the correlations in the 2022 CES survey data to the 2020 CES survey data, and to see if there are any significant correlations that can be used to predict a voter's alignment.
+Our main focus is to compare the correlations in the 2022 CES survey data to the 2020 CES survey data, and to see if our estimand, the effect of a voter's cultural background and political stances on their voting decision, has influenced their voting decision in the 2020 election.
 Using these understandings, we hope to highlight the potential implications these correlations may have for this upcoming 2024 election.
@@
 With the two main parties, Democrats and Republicans, having opposing stances on these issues, we found that there was a strong correlation between these stances and who a voter supported.
-As Biden was in favor of student loan forgiveness, but opposed to gun ownership, while Trump was opposed to student loan forgiveness, but in favor of gun ownership, we discovered a correlation between a voter's ties to these stances because they directly oppose each other.
+As Biden was in favour of student loan forgiveness, but opposed to gun ownership, while Trump was opposed to student loan forgiveness, but in favour of gun ownership, we discovered a correlation between a voter's ties to these stances because they directly oppose each other.
 This is just one example of the many correlations we found, and we will discuss these in more detail in the results section.
@@
 
+<!-- A sense of the dataset should be communicated to the reader.
+The broader context of the dataset should be discussed.
+All variables should be thoroughly examined and explained.
+Explain if there were similar datasets that could have been used and why they were not.
+If variables were constructed then this should be mentioned, and high-level cleaning aspects of note should be mentioned, but this section should focus on the destination, not the journey.
+It is important to understand what the variables look like by including graphs, and possibly tables, of all observations, along with discussion of those graphs and the other features of these data.
+Summary statistics should also be included, and well as any relationships between the variables.
+If this becomes too detailed, then appendices could be used.
+-->
+
 These delay statistics were downloaded, cleaned, parsed, analyzed, and visualized using R [@citeR], a statistical programming language, with package support from `tidyverse` [@citetidyverse], a collection of libraries which included the following packages that were utilized:
@@
 The data was retrieved from Harvard's CCES Dataverse database, using the `dataverse` [@citedataverse] package.
+In this report we consider data collected from the 2022 CES survey, stored by Dataverse as *Cooperative Election Study Common Content* [@citecesdata].
+There was an option to use the 2020 CES dataset, but we opted to use the newer, 60000 observation, 2022 CES dataset for a more updated and accurate representation of the current political climate in the United States after two years of Joe Biden being in office.
 
-In this report we consider data collected from the 2022 CES survey, stored by Dataverse as *Cooperative Election Study Common Content* [@citecesdata].
+Possible put in appendix?
+The variables we used in our analysis were: 
+- `presvote20post`, renamed to `voted_for`
+  - This data was represented in the form of a numerical variable corresponding to the presidential candidate that voted for in the 2020 eleciton.
+This was cleaned to limit the options to either "Biden" or "Trump", the two leading candidates in the 2020 and 2024 election.
+- `birthyr`, which was used to calculate the age of the voter, and later grouped into age buckets and renamed to `age`.
+  - This data was represented in the form of a numerical variable corresponding to the year the voter was born.
+This was cleaned to calculate the age of the voter, and then grouped into the age buckets: "18-29", "30-44", "45-64", "65+".
+- `gender4`, renamed to `gender`,
+  - This data was represented in the form of a numerical variable corresponding to the gender of the voter. This was cleaned to limit the options to "Man", "Woman", and "Other".
+- `educ`, renamed to education,
+  - This data was reprented in the form of a numerical variable corresponding to the highest level of education the voter has completed.
+This was cleaned to limit the options to "No HS" "High School graduate", "Some College", "2-year", "4-year", "Post-grad".
+- `race`, 
+  - This data was represented in the form of a numerical variable corresponding to the ethnicity of the voter.
+This was cleaned to limit the options to "White", "Black", "Hispanic", "Asian", "Middle Eastern", "Native American", "Multiracial", and "Other".
+- `religpew`, renamed to `religion`, 
+  - This data was represented in the form of a numerical variable corresponding to the religion of the voter.
+This was cleaned to limit the options to "Protestant", "Roman Catholic", "Jewish", "Atheist", "Agnostic", "Non-Religious", and "Other".
+There were various other religions that were not included in the dataset, but represented a small portion of the sample, and were thus grouped into "Other".
+- `gun_own`, renamed to `gun_ownership`,
+  - This data was represented in the form of a numerical variable corresponding to whether the voter or anyone in their household owned a gun.
+This was cleaned to limit the options to "Yes", "No", and "Unsure".
+- `edloan`, renamed to `student_loan`,
+  - This data was represented in the form of a numerical variable corresponding to whether the voter had student loans. This was cleaned to limit the options to "Yes", "No".
+
+The reason that we chose a large amount of variables/factors for our paper is so that we can get a bigger representation of the political climate.
+Should this be limited to a single model, then it may not reflect other correlations that may be present in the data.
+By using these variables, we take the most significant factors that may influence a voter's decision, and use them to predict who a voter supported in the 2020 election `voted_for`.
+It was important to get the voter's age, `age`, because it is important to understand the generational differences and how their experiences and personal interests may influence their voting decision.
+In our dataset, it was clear that 'Female' respondents were in more support towards the Democrats at 52% as opposed to 43% for 'Males'.
+In our given age buckets, it was discovered that 65% of 18-29 year olds supported the Democrats, whereas only 40% of 65+ year olds supported the Democrats, however, 18-29 year olds only made up 12% of the sample, whereas 65+ year olds made up 30% of the sample.
+It was also important to get the voter's gender, `gender`, and race, `race`, because it is important to understand the biological differences and how their experiences and personal interests may influence their voting decision.
+In our dataset, it was discovered that 84% and 71% of Black and Asian respondents, respectively, supported the Democrats, whereas only 41% of White respondents supported the Democrats.
+Incoorporating the voter's religion, `religion`, was important because it is important to understand that religious beliefs and differences can have a strong influence on a person's personal beliefs and therefore impact their personal stances on topics such as gun ownership.
+To put a contrasting example out there, 30% of Protestant/Christian respondents support the Democrats as opposed to a whopping 72% by Atheists.
+This could be attributed to certain parties aligning more to certain religious beliefs.
+A voter's education, `education`, was important because there may be certain understandings they arrive at during the completion of their education.
+Perhaps they viewed the world in one way at first, but after becoming more 'educated' in a topic, feel they have a better understanding of the world and therefore have a different stance on certain topics.
+This goes hand in hand with the voter's stance on student loans, `student_loan`, as it is important to understand how a voter's personal experiences with student loans may influence their voting decision.
+There was a time of serious consideration during Biden's tenure where he was considering forgiving student loans, and this intent alone may have swayed some support in his favour.
 
 needs measurement here
+<!-- Some aspect of measurement, relating to the dataset, is mentioned in the data section. -->
 

### 2024-03-14 scripts/02_clean_data.R

@@ -70,7 +70,7 @@ cleaned_ces2022 <-
     age_group = case_when(
       age < 30 ~ "18-29",
       age >= 30 & age < 45 ~ "30-44",
-      age >= 45 & age < 65 ~ "45-65",
+      age >= 45 & age < 65 ~ "45-64",
       age >= 65 ~ "65+"
     ),
     gender = case_when(

### 2024-03-14 output/paper.qmd
remove placeholders

@@
 
-<!-- A sense of the dataset should be communicated to the reader.
-The broader context of the dataset should be discussed.
-All variables should be thoroughly examined and explained.
-Explain if there were similar datasets that could have been used and why they were not.
-If variables were constructed then this should be mentioned, and high-level cleaning aspects of note should be mentioned, but this section should focus on the destination, not the journey.
-It is important to understand what the variables look like by including graphs, and possibly tables, of all observations, along with discussion of those graphs and the other features of these data.
-Summary statistics should also be included, and well as any relationships between the variables.
-If this becomes too detailed, then appendices could be used.
--->
-
 These delay statistics were downloaded, cleaned, parsed, analyzed, and visualized using R [@citeR], a statistical programming language, with package support from `tidyverse` [@citetidyverse], a collection of libraries which included the following packages that were utilized:
@@
 There was a time of serious consideration during Biden's tenure where he was considering forgiving student loans, and this intent alone may have swayed some support in his favour.
-
-needs measurement here
-<!-- Some aspect of measurement, relating to the dataset, is mentioned in the data section. -->
 

### 2024-03-15 output/paper.qmd
Tweak data section

@@
 
+For additional assitance with report generation the `knitr` [@citeknitr] package was used.
+
 The data was retrieved from Harvard's CCES Dataverse database, using the `dataverse` [@citedataverse] package.
-In this report we consider data collected from the 2022 CES survey, stored by Dataverse as *Cooperative Election Study Common Content* [@citecesdata].
+In this report we consider data collected from the 2022 CES (Cooperative ELection Study) survey, a questionnaire administered by YouGov collecting information about respondent's demographic, background, beliefs, opinions, past ballots cast, and future vote intentions.
+This data is accessible through Dataverse as *Cooperative Election Study Common Content* [@citecesdata].
 There was an option to use the 2020 CES dataset, but we opted to use the newer, 60000 observation, 2022 CES dataset for a more updated and accurate representation of the current political climate in the United States after two years of Joe Biden being in office.
+When the 2024 elections are held, we can see how the original previous aligning of the respondents have changed, if they have changed.
 
-Possible put in appendix?
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-cesdata
+#| tbl-cap: "2022 CES "
+
+# Load in data
+ces2022_1 <- read_csv("data/cleaned_ces2022_1.csv")
+ces2022_2 <- read_csv("data/cleaned_ces2022_2.csv")
+
+ces2022_2 |>
+  head() |>
+  kable(
+    col.names = c("Voted For (2020)", "Age Group", "Gender", "Education", "Race", "Religion", "Household Gun Ownership", "Student Loan Status"),
+    booktabs = TRUE
+  )
+```
+
 The variables we used in our analysis were: 
@@
 In our dataset, it was discovered that 84% and 71% of Black and Asian respondents, respectively, supported the Democrats, whereas only 41% of White respondents supported the Democrats.
-Incoorporating the voter's religion, `religion`, was important because it is important to understand that religious beliefs and differences can have a strong influence on a person's personal beliefs and therefore impact their personal stances on topics such as gun ownership.
+Incorporating the voter's religion, `religion`, was important because it is important to understand that religious beliefs and differences can have a strong influence on a person's personal beliefs and therefore impact their personal stances on topics such as gun ownership.
 To put a contrasting example out there, 30% of Protestant/Christian respondents support the Democrats as opposed to a whopping 72% by Atheists.
@@
 ```{r}
-#| include: false
+#| echo: false
+#| message: false
 #| warning: false
-#| message: false
-#| echo: false
+#| label: tbl-
+#| tbl-cap: Sample of Subway Delay Statistics
 
-# Load in data
-ces2022_1 <- read_csv("data/cleaned_ces2022_1.csv")
-ces2022_2 <- read_csv("data/cleaned_ces2022_2.csv")
+ces2022_1 |>
+  filter(gender != "Other", race %in%  c("Asian", "Black", "Hispanic")) |>
+  ggplot(aes(x = religion, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(race)) +
+  theme_minimal() +
+  labs(
+    x = "Race",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
+  theme(legend.position = "bottom")
 ```
@@
 ces2022_1 |>
-  ggplot(aes(x = education, fill = voted_for)) +
+  filter(race %in%  c("Asian", "Black", "Hispanic")) |>
+  ggplot(aes(x = religion, fill = voted_for)) +
   stat_count(position = "dodge") +
-  facet_wrap(facets = vars(gender)) +
+  facet_wrap(facets = vars(race)) +
   theme_minimal() +
   labs(
-    x = "Highest education",
+    x = "Race",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
+  theme(legend.position = "bottom")
+
+ces2022_1 |>
+  filter(!(race %in%  c("Asian", "Black", "Hispanic"))) |>
+  ggplot(aes(x = religion, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(race)) +
+  theme_minimal() +
+  labs(
+    x = "Race",
     y = "Number of respondents",
@@
 For the 2020 election, the candidates we are predicting for, Joe Biden and Donald Trump had opposing positions on student loan forgiveness.
-Whilst Trump was opposed to student loan forgiveness, backing incone-based repayment plans instead ()[@citetrumploanstance], Biden proposed to forgive a large amount of undergraduate tuition-related student debt ()[@citeballotpedia].
-They also opposed each other on gun laws, with Trump advocating for the right to bear arms and Biden advocating gun safety policies aimed at holding manufacturers accountable ()[@citeballotpedia].
+Whilst Trump proposed grants that would help with future students, Biden proposed to forgive a large amount of undergraduate tuition-related student debt [@citeballotpedia].
+In similar opposition, they held different opinions on gun laws, with Trump advocating for the right to bear arms and Biden advocating gun safety policies aimed at holding manufacturers accountable [@citeballotpedia].
 
 As such, we expect that there would be a strong correlation with the respondents gun ownership and student loan status to the candidate they supported in the 2020 election.
-
-TODO add to refs
-
-https://www.washingtonpost.com/politics/interactive/2023/presidential-candidates-2024-policies-issues/donald-trump-student-loans-education/
-
-https://ballotpedia.org/2020_presidential_candidates_on_student_loan_debt
-https://www.rand.org/research/gun-policy/key-findings/gun-policy-in-america.html
 
@@
 
+# Rename coefficients
+cm <- c('genderMale' = 'Male',
+        'genderOther' = 'Other Gender',
+        'raceOther' = 'Other Race',
+        'raceWhite' = 'White',
+        'raceNative American' = 'Native American', 
+        'raceMultiracial' = 'Multiracial',
+        'raceMiddle Eastern' = 'Middle Eastern',
+        'raceHispanic' = 'Hispanic',
+        'raceBlack' = 'Black',
+        'raceAsian' = 'Asian',
+        'religionNone' = 'Non-Religious',
+        'religionProtestant' = 'Protestant',
+        'religionAgnostic' = 'Agnostic',
+        'religionAtheist' = 'Atheist',
+        'religionRoman Catholic' = 'Roman Catholic',
+        'religionJewish' = 'Jewish',
+        'religionOther' = 'Other',
+        'age_group30-44' = 'Age (30-44)',
+        'age_group45-65' = 'Age (45-65)',
+        'age_group65+' = 'Age (65+)'
+        )
+
 modelsummary(
@@
   ),
-  statistic = "mad"
+  statistic = "mad",
+  coef_rename = cm
   )
@@
 ```{r}
+#| label: tbl-modelsummaryimportances
+#| tbl-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their education, household gun ownership, and student load status."
+#| message: false
+#| warning: false
+#| echo: false
+
+cmi <- c('education4-year' = '4 year Education',
+         'educationHigh school graduate' = 'High School',
+         'educationNo HS' = 'No High School',
+         'educationPost-grad' = 'Post Graduate',
+         'educationSome college' = 'Some College',
+         'household_gun_ownershipYes' = 'Household Owns Guns',
+         'student_loan_statusYes' = 'Has Student Loans'
+         )
+
 modelsummary(
@@
   ),
-  statistic = "mad"
+  statistic = "mad",
+  coef_rename = cmi
   )

### 2024-03-15 scripts/01_download_data.R
Update csv pathing in scripts

@@ -32,5 +32,5 @@ ces2022 <-
   # edloan        => If they have education loans
 
 # Save to csv
-write_csv(ces2022, "input/ces2022.csv")
+write_csv(ces2022, "input/data/ces2022.csv")
 

### 2024-03-15 scripts/02_clean_data.R

@@ -15,7 +15,7 @@ library(tidyverse)
 # Read in data
 ces2022 <- 
   read_csv(
-    "input/ces2022.csv",
+    "input/data/ces2022.csv",
     col_types =
       cols(
         "presvote20post" = col_integer(),
@@ -145,9 +145,11 @@ write_csv(cleaned_ces2022_1, "output/data/cleaned_ces2022_1.csv")
 cleaned_ces2022_2 <-
   cleaned_ces2022 |>
   filter(
-    !is.na(gunown),
-    !is.na(edloan)
+    !is.na(gunown)
   ) |>
+  filter(
+    !is.na(edloan)
+  )|>
   mutate(
     household_gun_ownership = case_when(
       gunown == 1 ~ "Yes",

### 2024-03-15 scripts/02_clean_data.R
Clean out of range gun own values

@@ -155,7 +155,8 @@ cleaned_ces2022_2 <-
       gunown == 1 ~ "Yes",
       gunown == 2 ~ "Yes",
       gunown == 3 ~ "No",
-      gunown == 4 ~ "Unsure"
+      gunown == 4 ~ "Unsure",
+      .default = "Unsure"
     ),
     student_loan_status = if_else(
       edloan == 1,

### 2024-03-15 output/paper.qmd
Add graphics
NOT RENDERING

@@
 bibliography: references.bib
-editor: visual
+header-includes:
+  \usepackage{placeins}
 ---
@@
 The data was retrieved from Harvard's CCES Dataverse database, using the `dataverse` [@citedataverse] package.
-In this report we consider data collected from the 2022 CES (Cooperative ELection Study) survey, a questionnaire administered by YouGov collecting information about respondent's demographic, background, beliefs, opinions, past ballots cast, and future vote intentions.
+In this report we consider data collected from the 2022 CES (Cooperative Election Study) survey, a questionnaire administered by YouGov collecting information about respondent's demographic, background, beliefs, opinions, past ballots cast, and future vote intentions.
 This data is accessible through Dataverse as *Cooperative Election Study Common Content* [@citecesdata].
@@
 #| label: tbl-cesdata
-#| tbl-cap: "2022 CES "
+#| tbl-cap: "2022 CES (Cooperative Election Study) Data"
 
@@
 ```{r}
+#| eval: true
 #| echo: false
@@
 #| warning: false
-#| label: tbl-
-#| tbl-cap: Sample of Subway Delay Statistics
+#| fig-cap: "Votes for Candidate given Age Group and Religion"
+#| label: fig-agevsreligion
 
 ces2022_1 |>
-  filter(gender != "Other", race %in%  c("Asian", "Black", "Hispanic")) |>
   ggplot(aes(x = religion, fill = voted_for)) +
   stat_count(position = "dodge") +
-  facet_wrap(facets = vars(race)) +
+  facet_wrap(facets = vars(age_group)) +
   theme_minimal() +
   labs(
-    x = "Race",
-    y = "Number of respondents",
-    fill = "Voted for"
-  ) +
-  coord_flip() +
-  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
-  theme(legend.position = "bottom")
-```
-
-```{r}
-ces2022_1 |>
-  filter(race %in%  c("Asian", "Black", "Hispanic")) |>
-  ggplot(aes(x = religion, fill = voted_for)) +
-  stat_count(position = "dodge") +
-  facet_wrap(facets = vars(race)) +
-  theme_minimal() +
-  labs(
-    x = "Race",
+    x = "Religion",
     y = "Number of respondents",
@@
 
-ces2022_1 |>
-  filter(!(race %in%  c("Asian", "Black", "Hispanic"))) |>
-  ggplot(aes(x = religion, fill = voted_for)) +
+```
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-cap: "Votes for Candidate given Gun Ownership and Education"
+#| label: tbl-gunownvseducation
+
+ces2022_2 |>
+  ggplot(aes(x = household_gun_ownership, fill = voted_for)) +
   stat_count(position = "dodge") +
-  facet_wrap(facets = vars(race)) +
+  facet_wrap(facets = vars(education)) +
   theme_minimal() +
   labs(
-    x = "Race",
+    x = "Does someone in your household own a gun?",
     y = "Number of respondents",
@@
 ```{r}
+#| include: false
+#| warning: false
+#| message: false
+#| echo: false
+
 # Load in models
@@
 
-political_preferences_race <-
-  readRDS(file = "models/political_preferences_race.rds")
-
 political_preferences_importances <-
   readRDS(file = "models/political_preferences_importances.rds")
+
 ```
@@
 
-We expect that due to various differences in experiences due to this background, whether it be because of a generational difference, a difference in culture between countries, or a difference in beliefs in religions, a voter who identifies in those demographics would vote differently, voting for the candidate who most undestands their experiences.
+We expect that due to various differences in experiences due to this background, whether it be because of a generational difference, a difference in culture between countries, or a difference in beliefs in religions, a voter who identifies in those demographics would vote differently, voting for the candidate who most understands their experiences.
 
-The second model only forecasts who the respondent was likely to vote for based on race, to extract how much of the intercept in the previous model can be explained by a different in race.
-
-Finally, the last model aims to forecast a respondents vote knowing their prominent political ideals including education, gun ownership status, and student loan status.
+The second model aims to forecast a respondents vote knowing their prominent political ideals including education, gun ownership status, and student loan status.
 These are common political standpoints in the U.S..
@@
 ```{r}
-#| label: tbl-modelsummarycultural
-#| tbl-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their gender, age, race, and religion"
+#| label: fig-modelsummarycultural
+#| fig-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their gender, age, race, and religion"
 #| message: false
@@
 ```{r}
-#| label: tbl-modelsummaryimportances
-#| tbl-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their education, household gun ownership, and student load status."
+#| label: fig-modelsummaryimportances
+#| fig-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their education, household gun ownership, and student load status."
 #| message: false
@@
 ```{r}
-cm <- c('genderMale' = 'Male',
-        'raceWhite' = 'White',
-        'raceNative American' = 'Native American', 
-        'raceMultiracial' = 'Multiracial',
-        'raceMiddle Eastern' = 'Middle Eastern',
-        'raceHispanic' = 'Hispanic',
-        'raceBlack' = 'Black',
-        'raceAsian' = 'Asian',
-        'religionNone' = 'Non-Religious',
-        'religionProtestant' = 'Protestant',
-        'religionAgnostic' = 'Agnostic',
-        'religionAtheist' = 'Atheist',
-        'religionRoman Catholic' = 'Roman Catholic',
-        'religionJewish' = 'Jewish',
-        'religionOther' = 'Other'
-        )
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-conf_cultural
+#| tbl-cap: "Cultural Model 90% Confidence Intervals"
 
@@
   labs(x = "90 per cent credibility interval")
+
 ```
@@
 ```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-confother
+#| fig-cap: "Cultural Model 90% Confidence Intervals (Race and Gender Other)"
+
 modelplot(political_preferences_cultural, 
@@
   labs(x = "90 per cent credibility interval")
+
 ```
 
-\newpage
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-confimportances
+#| fig-cap: "Key Topics Model 90% Confidence Intervals"
+
+modelplot(political_preferences_importances, conf_level = 0.9, coef_map = cmi) +
+  labs(x = "90 per cent credibility interval")
+
+```
+
 ## References

### 2024-03-15 output/paper.qmd
Change figs to tables

@@
 ```{r}
-#| label: fig-modelsummarycultural
-#| fig-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their gender, age, race, and religion"
 #| message: false
@@
 #| echo: false
+#| label: tbl-modelsummarycultural
+#| tbl-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their gender, age, race, and religion"
 
@@
 ```{r}
-#| label: fig-modelsummaryimportances
-#| fig-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their education, household gun ownership, and student load status."
 #| message: false
@@
 #| echo: false
+#| label: tbl-modelsummaryimportances
+#| tbl-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their education, household gun ownership, and student load status."
 

### 2024-03-15 output/paper.qmd
Clean up graphs

@@
 We will also discuss the implications of these findings, and how they may be used to predict a voter's alignment in the upcoming 2024 election.
-
-TODO: cite https://www.theguardian.com/us-news/2020/nov/07/joe-biden-wins-us-election-donald-trump-loses-final-result-2020
 
@@
 #| fig-cap: "Votes for Candidate given Gun Ownership and Education"
-#| label: tbl-gunownvseducation
+#| label: fig-gunownvseducation
 
@@
 
-## Results
+# Results
 
-## Discussion
+# Discussion
 
@@
 
-## Appendix
+# Appendix
 
@@
 #| echo: false
-#| label: tbl-conf_cultural
-#| tbl-cap: "Cultural Model 90% Confidence Intervals"
+#| label: fig-conf_cultural
+#| fig-cap: "Cultural Model 90% Confidence Intervals"
+
+cm <- c('genderMale' = 'Male',
+        'raceWhite' = 'White',
+        'raceNative American' = 'Native American', 
+        'raceMultiracial' = 'Multiracial',
+        'raceMiddle Eastern' = 'Middle Eastern',
+        'raceHispanic' = 'Hispanic',
+        'raceBlack' = 'Black',
+        'raceAsian' = 'Asian',
+        'religionNone' = 'Non-Religious',
+        'religionProtestant' = 'Protestant',
+        'religionAgnostic' = 'Agnostic',
+        'religionAtheist' = 'Atheist',
+        'religionRoman Catholic' = 'Roman Catholic',
+        'religionJewish' = 'Jewish',
+        'religionOther' = 'Other',
+        'age_group30-44' = 'Age (30-44)',
+        'age_group45-65' = 'Age (45-65)',
+        'age_group65+' = 'Age (65+)'
+        )
 
@@
           conf_level = 0.9, 
-          coef_map = c("raceOther" = "Other Race", "genderOther" = "Other Gender (Does not Identify as Male or Female)")) + 
+          coef_map = c("raceOther" = "Other Race", "genderOther" = "Other Gender")) + 
   labs(x = "90 per cent credibility interval")
@@
 
-## References
+\newpage
+# References

### 2024-03-16 output/paper.qmd
split tables

@@
 #| echo: false
-#| label: tbl-cesdata
-#| tbl-cap: "2022 CES (Cooperative Election Study) Data"
+#| label: tbl-cesdata1
+#| tbl-cap: "2022 CES (Cooperative Election Study) Data (Cultural)"
 
@@
 
-ces2022_2 |>
+ces2022_1 |>
   head() |>
   kable(
-    col.names = c("Voted For (2020)", "Age Group", "Gender", "Education", "Race", "Religion", "Household Gun Ownership", "Student Loan Status"),
+    col.names = c("Voted For (2020)", "Age Group", "Gender", "Education", "Race", "Religion"),
     booktabs = TRUE
   )
+```
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-cesdata2
+#| tbl-cap: "2022 CES (Cooperative Election Study) Data (Cultural) Standings"
+
+ces2022_2 |>
+  select(voted_for, household_gun_ownership, student_loan_status) |>
+  head() |>
+  kable(
+    col.names = c("Voted For (2020)", "Household Gun Ownership", "Student Loan Status"),
+    booktabs = TRUE
+  )
+
 ```
@@
 
-# Appendix
+# Appendix {-}
 

### 2024-03-16 output/paper.qmd
formatted data section

@@
 
-The variables we used in our analysis were: 
-- `presvote20post`, renamed to `voted_for`
-  - This data was represented in the form of a numerical variable corresponding to the presidential candidate that voted for in the 2020 eleciton.
+The variables we used in our analysis were: - `presvote20post`, renamed to `voted_for` - This data was represented in the form of a numerical variable corresponding to the presidential candidate that voted for in the 2020 eleciton.
 This was cleaned to limit the options to either "Biden" or "Trump", the two leading candidates in the 2020 and 2024 election.
+
 -   `birthyr`, which was used to calculate the age of the voter, and later grouped into age buckets and renamed to `age`.
+
     -   This data was represented in the form of a numerical variable corresponding to the year the voter was born.
 This was cleaned to calculate the age of the voter, and then grouped into the age buckets: "18-29", "30-44", "45-64", "65+".
+
 -   `gender4`, renamed to `gender`,
+
     -   This data was represented in the form of a numerical variable corresponding to the gender of the voter. This was cleaned to limit the options to "Man", "Woman", and "Other".
+
 -   `educ`, renamed to education,
+
     -   This data was reprented in the form of a numerical variable corresponding to the highest level of education the voter has completed.
 This was cleaned to limit the options to "No HS" "High School graduate", "Some College", "2-year", "4-year", "Post-grad".
+
 -   `race`,
+
     -   This data was represented in the form of a numerical variable corresponding to the ethnicity of the voter.
 This was cleaned to limit the options to "White", "Black", "Hispanic", "Asian", "Middle Eastern", "Native American", "Multiracial", and "Other".
+
 -   `religpew`, renamed to `religion`,
+
     -   This data was represented in the form of a numerical variable corresponding to the religion of the voter.
@@
 There were various other religions that were not included in the dataset, but represented a small portion of the sample, and were thus grouped into "Other".
+
 -   `gun_own`, renamed to `gun_ownership`,
+
     -   This data was represented in the form of a numerical variable corresponding to whether the voter or anyone in their household owned a gun.
 This was cleaned to limit the options to "Yes", "No", and "Unsure".
+
 -   `edloan`, renamed to `student_loan`,
+
     -   This data was represented in the form of a numerical variable corresponding to whether the voter had student loans. This was cleaned to limit the options to "Yes", "No".
@@
 By using these variables, we take the most significant factors that may influence a voter's decision, and use them to predict who a voter supported in the 2020 election `voted_for`.
-It was important to get the voter's age, `age`, because it is important to understand the generational differences and how their experiences and personal interests may influence their voting decision.
-In our dataset, it was clear that 'Female' respondents were in more support towards the Democrats at 52% as opposed to 43% for 'Males'.
+
+    It was important to get the voter's `age`, in order to understand the generational differences and how their experiences and personal interests may influence their voting decision.
 In our given age buckets, it was discovered that 65% of 18-29 year olds supported the Democrats, whereas only 40% of 65+ year olds supported the Democrats, however, 18-29 year olds only made up 12% of the sample, whereas 65+ year olds made up 30% of the sample.
-It was also important to get the voter's gender, `gender`, and race, `race`, because it is important to understand the biological differences and how their experiences and personal interests may influence their voting decision.
-In our dataset, it was discovered that 84% and 71% of Black and Asian respondents, respectively, supported the Democrats, whereas only 41% of White respondents supported the Democrats.
-Incorporating the voter's religion, `religion`, was important because it is important to understand that religious beliefs and differences can have a strong influence on a person's personal beliefs and therefore impact their personal stances on topics such as gun ownership.
+
+    It was also necessary to get the voter's `gender`, and `race`, because it is important to understand the biological differences and how their experiences and personal interests may influence their voting decision.
+In our dataset, it was clear that 'Female' respondents were more supportive towards the Democrats at 52% as opposed to 43% for 'Males'.
+In the racial side of the respondent data, it was discovered that 84% and 71% of Black and Asian respondents, respectively, supported the Democrats, whereas only 41% of White respondents supported the Democrats.
+
+    Incorporating the voter's `religion`, was a needed factor since it is important to understand that religious beliefs and differences can have a strong influence one's personal beliefs and therefore impact their personal stances on topics such as gun ownership.
 To put a contrasting example out there, 30% of Protestant/Christian respondents support the Democrats as opposed to a whopping 72% by Atheists.
 This could be attributed to certain parties aligning more to certain religious beliefs.
-A voter's education, `education`, was important because there may be certain understandings they arrive at during the completion of their education.
-Perhaps they viewed the world in one way at first, but after becoming more 'educated' in a topic, feel they have a better understanding of the world and therefore have a different stance on certain topics.
+
+    A voter's `education`, was important because there may be certain understandings they arrive at during the completion of their education.
+Perhaps they viewed the world in one way at first, but after becoming more 'educated' in a topic, they feel that they have a better understanding of the world and either become more confident in their initial understanding or change their stance entirely.
 This goes hand in hand with the voter's stance on student loans, `student_loan`, as it is important to understand how a voter's personal experiences with student loans may influence their voting decision.

### 2024-03-16 output/paper.qmd
Add all variable plots

@@
 
-These delay statistics were downloaded, cleaned, parsed, analyzed, and visualized using R [@citeR], a statistical programming language, with package support from `tidyverse` [@citetidyverse], a collection of libraries which included the following packages that were utilized:
+The data was retrieved from Harvard's CCES Dataverse database, using the `dataverse` [@citedataverse] package.
+In this report we consider data collected from the 2022 CES (Cooperative Election Study) survey, a questionnaire administered by YouGov collecting information about respondent's demographic, background, beliefs, opinions, past ballots cast, and future vote intentions.
+This data is accessible through Dataverse as *Cooperative Election Study Common Content* [@citecesdata].
+There was an option to use the 2020 CES dataset, but we opted to use the newer, 60000 observation, 2022 CES dataset for a more updated and accurate representation of the current political climate in the United States after two years of Joe Biden being in office.
+When the 2024 elections are held, we can see how the original previous aligning of the respondents have changed, if they have changed.
+
+This data was downloaded, cleaned, parsed, analyzed, and visualized using R [@citeR], a statistical programming language, with package support from `tidyverse` [@citetidyverse], a collection of libraries which included the following packages that were utilized:
 
@@
 For additional assitance with report generation the `knitr` [@citeknitr] package was used.
-
-The data was retrieved from Harvard's CCES Dataverse database, using the `dataverse` [@citedataverse] package.
-In this report we consider data collected from the 2022 CES (Cooperative Election Study) survey, a questionnaire administered by YouGov collecting information about respondent's demographic, background, beliefs, opinions, past ballots cast, and future vote intentions.
-This data is accessible through Dataverse as *Cooperative Election Study Common Content* [@citecesdata].
-There was an option to use the 2020 CES dataset, but we opted to use the newer, 60000 observation, 2022 CES dataset for a more updated and accurate representation of the current political climate in the United States after two years of Joe Biden being in office.
-When the 2024 elections are held, we can see how the original previous aligning of the respondents have changed, if they have changed.
 
@@
 #| warning: false
-#| fig-cap: "Votes for Candidate given Age Group and Religion"
+#| fig-cap: "Votes for Candidate given Age Group and Gender"
+#| fig-subcap: ["Male and Female", "Other (Non-binary and Other)"]
 #| label: fig-agevsreligion
+#| layout-ncol: 2
 
 ces2022_1 |>
-  ggplot(aes(x = religion, fill = voted_for)) +
+  filter(gender != "Other") |>
+  ggplot(aes(x = age_group, fill = voted_for)) +
   stat_count(position = "dodge") +
-  facet_wrap(facets = vars(age_group)) +
+  facet_wrap(facets = vars(gender)) +
   theme_minimal() +
   labs(
-    x = "Religion",
+    x = "Age Group",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
+  theme(legend.position = "bottom")
+
+ces2022_1 |>
+  filter(gender == "Other") |>
+  ggplot(aes(x = age_group, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender)) +
+  theme_minimal() +
+  labs(
+    x = "Age Group",
     y = "Number of respondents",
@@
   labs(
-    x = "Does someone in your household own a gun?",
+    x = "Household Gun Ownership",
     y = "Number of respondents",
@@
   scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom", axis.text.x=element_text(angle=60, hjust=1))
+
+```
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-cap: "Votes for Candidate given Student Loan Status and Education"
+#| label: fig-loanvseducation
+
+ces2022_2 |>
+  ggplot(aes(x = education, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(student_loan_status)) +
+  theme_minimal() +
+  labs(
+    x = "Education",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
+  theme(legend.position = "bottom", axis.text.x=element_text(angle=60, hjust=1))
+
+```
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-cap: "Votes for Candidate given Race and Religion"
+#| label: fig-racevsreligion1
+
+ces2022_2 |>
+  filter(race %in% c("Black", "Hispanic", "White")) |>
+  ggplot(aes(x = religion, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(race)) +
+  theme_minimal() +
+  labs(
+    x = "Religion",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
+  theme(legend.position = "bottom", axis.text.x=element_text(angle=60, hjust=1))
+
+```
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-cap: "Votes for Candidate given Race and Religion"
+#| label: fig-racevsreligion2
+
+ces2022_2 |>
+  filter(!(race %in% c("Black", "Hispanic", "White"))) |>
+  ggplot(aes(x = religion, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(race)) +
+  theme_minimal() +
+  labs(
+    x = "Religion",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
+  theme(legend.position = "bottom", axis.text.x=element_text(angle=60, hjust=1))
 

### 2024-03-16 output/paper.qmd
Add model math

@@
 
-The second model aims to forecast a respondents vote knowing their prominent political ideals including education, gun ownership status, and student loan status.
-These are common political standpoints in the U.S..
-Gun policy and student loan forgiveness is often a key political issue that separates the candidates.
-For the 2020 election, the candidates we are predicting for, Joe Biden and Donald Trump had opposing positions on student loan forgiveness.
-Whilst Trump proposed grants that would help with future students, Biden proposed to forgive a large amount of undergraduate tuition-related student debt [@citeballotpedia].
-In similar opposition, they held different opinions on gun laws, with Trump advocating for the right to bear arms and Biden advocating gun safety policies aimed at holding manufacturers accountable [@citeballotpedia].
+Therefore the model we are interested in is:
 
-As such, we expect that there would be a strong correlation with the respondents gun ownership and student loan status to the candidate they supported in the 2020 election.
+$$
+\begin{aligned}
+y_i|\pi_i & \sim \mbox{Bern}(\pi_i) \\
+\mbox{logit}(\pi_i) & = \beta_0 +\beta_1 \times \mbox{age_group}_i + \beta_2 \times \mbox{gender}_i +\beta_3 \times \mbox{race}_i + \beta_4 \times \mbox{religion}_i \\
+\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_2 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_3 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_4 & \sim \mbox{Normal}(0, 2.5)
+\end{aligned}
+$$
+Where $y_i$ is whether they voted for Biden or not in the 2020 election, $\mbox{age_group}_i$ is the age group the respondent falls in, $\mbox{gender}_i$ is the gender of the respondent, $\mbox{race}_i$ is the race of the respondent, and $\mbox{religion}_i$ is the relgion of the respondent.
+
+@tbl-modelsummarycultural shows a model summary for this model, this will be dicussed further in the [results](#results) section.
 
@@
 
+The second model aims to forecast a respondents vote knowing their prominent political ideals including education, gun ownership status, and student loan status.
+These are common political standpoints in the U.S..
+Gun policy and student loan forgiveness is often a key political issue that separates the candidates.
+For the 2020 election, the candidates we are predicting for, Joe Biden and Donald Trump had opposing positions on student loan forgiveness.
+Whilst Trump proposed grants that would help with future students, Biden proposed to forgive a large amount of undergraduate tuition-related student debt [@citeballotpedia].
+In similar opposition, they held different opinions on gun laws, with Trump advocating for the right to bear arms and Biden advocating gun safety policies aimed at holding manufacturers accountable [@citeballotpedia].
+
+As such, we expect that there would be a strong correlation with the respondents gun ownership and student loan status to the candidate they supported in the 2020 election.
+
+$$
+\begin{aligned}
+y_i|\pi_i & \sim \mbox{Bern}(\pi_i) \\
+\mbox{logit}(\pi_i) & = \beta_0 +\beta_1 \times \mbox{education}_i + \beta_2 \times \mbox{student_loan_status}_i +\beta_3 \times \mbox{household_gun_ownership}_i \\
+\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_2 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_3 & \sim \mbox{Normal}(0, 2.5)
+\end{aligned}
+$$
+
+where $y_i$ is whether they voted for Biden or not in the 2020 election, $\mbox{education}_i$ is the highest education the respondent completed, $\mbox{student_loan_status}_i$ is if the respondent has student loans, and $\mbox{household_gun_ownership}_i$ is if the respondent or someone within their household owns a gun.
+
+@tbl-modelsummaryimportances shows a model summary for this model, this will be dicussed further in the [results](#results) section.
+
 ```{r}
@@
 
-# Results
+# Results {#results}
 

### 2024-03-16 output/paper.qmd
data update w/ cross ref, paper.pdf is outdated, need to fix latex $$ issue

@@
 
-  The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump (Gambino 2020).
+The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump [cite2020Election].
 Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors.
@@
 
-* `ggplot2` [@citeggplot] 
-* `dplyr` [@citedplyr] 
-* `readr` [@citereadr] 
-* `tibble` [@citetibble]
+-   `ggplot2` [@citeggplot]
+-   `dplyr` [@citedplyr]
+-   `readr` [@citereadr]
+-   `tibble` [@citetibble]
 
@@
 
+Personal Background:
+
 -   `birthyr`, which was used to calculate the age of the voter, and later grouped into age buckets and renamed to `age`.
@@
 
--   `gun_own`, renamed to `gun_ownership`,
+    A small sample of these variables can be seen in the @tbl-cesdata1, which shows rows of respondent data, including their `voted_for`, `age_group`, `gender`, `education`, `race`, and `religion`.
+
+Personal Stances:
+
+-   `gun_own`, renamed to `household_gun_ownership`,
 
@@
 
+    A small sample of these variables can be seen in the @tbl-cesdata2, which shows rows of respondent data, including their `voted_for`, `household_gun_ownership`, and `student_loan_status`.
+
     The reason that we chose a large amount of variables/factors for our paper is so that we can get a bigger representation of the political climate.
@@
 In our given age buckets, it was discovered that 65% of 18-29 year olds supported the Democrats, whereas only 40% of 65+ year olds supported the Democrats, however, 18-29 year olds only made up 12% of the sample, whereas 65+ year olds made up 30% of the sample.
+In @fig-agevsgender, we can see how respondents' support for the Democrats and Republicans can differ based on their `age` and `gender`.
+For example, in @fig-agevsgender-1 'Males' aged 65+ were quite a bit more supportive towards the Republicans as opposed to 'Males' aged 18-29, who showed a noticibly larger gap in support towards the Democrats.
+In @fig-agevsgender-2, despite the lower respondent count, we can see how 'Other' gendered respondents had significantly more voters in the 18-29, 30-44 age groups, while also showing a larger gap in support towards the Democrats.
 
@@
     Incorporating the voter's `religion`, was a needed factor since it is important to understand that religious beliefs and differences can have a strong influence one's personal beliefs and therefore impact their personal stances on topics such as gun ownership.
-To put a contrasting example out there, 30% of Protestant/Christian respondents support the Democrats as opposed to a whopping 72% by Atheists.
+To put a contrasting example out there, 30% of Protestant/Christian respondents support the Democrats as opposed to 72% by Atheists.
 This could be attributed to certain parties aligning more to certain religious beliefs.
+To see a bigger insight into this, take a look at @fig-racevsreligion1 and @fig-racevsreligion2 (figures separated according to respondent count), and notice examples like, 'White Protestants' appear significantly more supportive towards the Republicans as opposed to 'White Catholics' and 'Black Protestants' who appear to be significantly more supportive towards the Democrats.
 
@@
 Perhaps they viewed the world in one way at first, but after becoming more 'educated' in a topic, they feel that they have a better understanding of the world and either become more confident in their initial understanding or change their stance entirely.
-This goes hand in hand with the voter's stance on student loans, `student_loan`, as it is important to understand how a voter's personal experiences with student loans may influence their voting decision.
-There was a time of serious consideration during Biden's tenure where he was considering forgiving student loans, and this intent alone may have swayed some support in his favour.
+@fig-gunownvseducation shows how `education` goes hand in hand with the voter's stance on gun ownership, `household_gun_ownership`, while @fig-loanvseducation shows how `education` goes hand in hand with the voter's stance on student loans, `student_loan`.
+It is important to understand how a voter's personal experiences with gun ownership and student loans may influence their voting decision.
+To highlight a key example, there was a time of serious consideration during Biden's tenure where he was considering forgiving student loans [@citeballotpedia], and this intent alone may have swayed some support in his favour.
 
@@
 #| fig-subcap: ["Male and Female", "Other (Non-binary and Other)"]
-#| label: fig-agevsreligion
+#| label: fig-agevsgender
 #| layout-ncol: 2
@@
 ```
+
+\newpage
 
@@
 
-Although there seems to be a strong correlation between if a voter
-identifies as an other gender to supporting Biden, and a strong correlation
-between if a voter identifies as an other race to supporting Biden,
-but these must be considered carefully as within the sample (with seed 302), 
-only 5 and 1 voters identified in those categories respectively. To get a more
-accurate reading in regards to these features, it would be necessary to train
-a model which includes much more of these demographics.
+Although there seems to be a strong correlation between if a voter identifies as an other gender to supporting Biden, and a strong correlation between if a voter identifies as an other race to supporting Biden, but these must be considered carefully as within the sample (with seed 302), only 5 and 1 voters identified in those categories respectively.
+To get a more accurate reading in regards to these features, it would be necessary to train a model which includes much more of these demographics.
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
@@
 \newpage
+
 # References

### 2024-03-16 output/paper.qmd
Add model math

@@
 library(rstanarm)
-
 ```
@@
 y_i|\pi_i & \sim \mbox{Bern}(\pi_i) \\
-\mbox{logit}(\pi_i) & = \beta_0 +\beta_1 \times \mbox{age_group}_i + \beta_2 \times \mbox{gender}_i +\beta_3 \times \mbox{race}_i + \beta_4 \times \mbox{religion}_i \\
+\mbox{logit}(\pi_i) & = \beta_0+\beta_1 \times \mbox{age-group}_i + \beta_2 \times \mbox{gender}_i \\
+&+ \beta_3 \times \mbox{race}_i + \beta_3 \times \mbox{religion}_i\\
 \beta_0 & \sim \mbox{Normal}(0, 2.5)\\
@@
 \beta_2 & \sim \mbox{Normal}(0, 2.5)\\
-\beta_3 & \sim \mbox{Normal}(0, 2.5)\\
-\beta_4 & \sim \mbox{Normal}(0, 2.5)
+\beta_3 & \sim \mbox{Normal}(0, 2.5)
 \end{aligned}
 $$
-Where $y_i$ is whether they voted for Biden or not in the 2020 election, $\mbox{age_group}_i$ is the age group the respondent falls in, $\mbox{gender}_i$ is the gender of the respondent, $\mbox{race}_i$ is the race of the respondent, and $\mbox{religion}_i$ is the relgion of the respondent.
 
-@tbl-modelsummarycultural shows a model summary for this model, this will be dicussed further in the [results](#results) section.
+Where $y_i$ is whether they voted for Biden or not in the 2020 election, $\text{age-group}_i$ is the age group the respondent falls in, $\text{gender}_i$ is the gender of the respondent, $\text{race}_i$ is the race of the respondent, and $\text{religion}_i$ is the religion of the respondent.
+
+@tbl-modelsummarycultural shows a model summary for this model, this will be discussed further in the [results](#results) section.
 
@@
 
+Therefore the model we are interested in is:
+
 $$
@@
 y_i|\pi_i & \sim \mbox{Bern}(\pi_i) \\
-\mbox{logit}(\pi_i) & = \beta_0 +\beta_1 \times \mbox{education}_i + \beta_2 \times \mbox{student_loan_status}_i +\beta_3 \times \mbox{household_gun_ownership}_i \\
+\mbox{logit}(\pi_i) & = \beta_0+\beta_1 \times \mbox{education}_i + \beta_2 \times \mbox{student-loan-status}_i \\
+& + \beta_3 \times \mbox{household-gun-ownership}_i\\
 \beta_0 & \sim \mbox{Normal}(0, 2.5)\\
@@
 
-where $y_i$ is whether they voted for Biden or not in the 2020 election, $\mbox{education}_i$ is the highest education the respondent completed, $\mbox{student_loan_status}_i$ is if the respondent has student loans, and $\mbox{household_gun_ownership}_i$ is if the respondent or someone within their household owns a gun.
+where $y_i$ is whether they voted for Biden or not in the 2020 election, $\text{education}_i$ is the highest education the respondent completed, $\text{student-loan-status}_i$ is if the respondent has student loans, and $\text{household-gun-ownership}_i$ is if the respondent or someone within their household owns a gun.
 
-@tbl-modelsummaryimportances shows a model summary for this model, this will be dicussed further in the [results](#results) section.
+@tbl-modelsummaryimportances shows a model summary for this model, this will be discussed further in the [results](#results) section.
 
@@
 # Results {#results}
+
+From those models we are able to see correlation between the factors chosen and their vote during the 2020 Presidential election.
+
+@tbl-modelsummarycultural
 

### 2024-03-16 scripts/03_model.R
Add comments

@@ -26,6 +26,7 @@ ces2022_1_reduced <-
   ces2022_1 |>
   slice_sample(n = 1000) |>
   mutate(
+    # Ensure that Biden is being selected as 1, and Trump is 0
     voted_for = if_else(voted_for == "Biden", 1, 0)
     # voted_for = as_factor(voted_for)
   )
@@ -34,6 +35,7 @@ ces2022_2_reduced <-
   ces2022_2 |>
   slice_sample(n = 1000) |>
   mutate(
+    # Ensure that Biden is being selected as 1, and Trump is 0
     voted_for = if_else(voted_for == "Biden", 1, 0)
     # voted_for = as_factor(voted_for)
   )
@@ -55,7 +57,7 @@ saveRDS(
   file = "output/models/political_preferences_cultural.rds"
 )
 
-# Model with only race
+# Model with only race, used to verify race's effect on the intercept.
 political_preferences_race <-
   stan_glm(
     voted_for ~ race,

### 2024-03-16 output/paper.qmd
Format sections and complete results

@@
 
-# Data
+# Data {#data}
 
@@
 
+## Variables
+
 The variables we used in our analysis were: - `presvote20post`, renamed to `voted_for` - This data was represented in the form of a numerical variable corresponding to the presidential candidate that voted for in the 2020 eleciton.
@@
 
-    -   This data was represented in the form of a numerical variable corresponding to the gender of the voter. This was cleaned to limit the options to "Man", "Woman", and "Other".
+    -   This data was represented in the form of a numerical variable corresponding to the gender of the voter.
+This was cleaned to limit the options to "Man", "Woman", and "Other".
+Other includes both respondents who specifically selected "Other" as well as respondents who responded as "Non-binary".
+Non-binary responses only totaled 354 participants out of 60,000.
 
@@
     -   This data was represented in the form of a numerical variable corresponding to the ethnicity of the voter.
-This was cleaned to limit the options to "White", "Black", "Hispanic", "Asian", "Middle Eastern", "Native American", "Multiracial", and "Other".
+This was cleaned to limit the options to "White", "Black", "Hispanic", "Asian", "Middle Eastern", "Native American", "Multiracial" and "Other".
 
@@
 This was cleaned to limit the options to "Protestant", "Roman Catholic", "Jewish", "Atheist", "Agnostic", "Non-Religious", and "Other".
-There were various other religions that were not included in the dataset, but represented a small portion of the sample, and were thus grouped into "Other".
+There were various other religions that were included in the dataset, but represented a small portion of the sample, and were thus grouped into "Other".
 
@@
     A small sample of these variables can be seen in the @tbl-cesdata2, which shows rows of respondent data, including their `voted_for`, `household_gun_ownership`, and `student_loan_status`.
+
+## Justification
 
@@
 This could be attributed to certain parties aligning more to certain religious beliefs.
-To see a bigger insight into this, take a look at @fig-racevsreligion1 and @fig-racevsreligion2 (figures separated according to respondent count), and notice examples like, 'White Protestants' appear significantly more supportive towards the Republicans as opposed to 'White Catholics' and 'Black Protestants' who appear to be significantly more supportive towards the Democrats.
+To see a bigger insight into this, take a look at @fig-racevsreligionlarge and @fig-racevsreligion2 (figures separated according to respondent count), and notice examples like, 'White Protestants' appear significantly more supportive towards the Republicans as opposed to 'White Catholics' and 'Black Protestants' who appear to be significantly more supportive towards the Democrats.
 
@@
 #| fig-cap: "Votes for Candidate given Race and Religion"
-#| label: fig-racevsreligion1
+#| label: fig-racevsreligionlarge
+#| layout-ncol: 2
 
 ces2022_2 |>
-  filter(race %in% c("Black", "Hispanic", "White")) |>
+  filter(race %in% c("Black", "Hispanic")) |>
+  ggplot(aes(x = religion, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(race)) +
+  theme_minimal() +
+  labs(
+    x = "Religion",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_fill_manual(values = c("Biden" = "blue", "Trump" = "red")) +
+  theme(legend.position = "bottom", axis.text.x=element_text(angle=60, hjust=1))
+
+ces2022_2 |>
+  filter(race %in% c("White")) |>
   ggplot(aes(x = religion, fill = voted_for)) +
@@
 
+political_preferences_race <-
+  readRDS(file = "models/political_preferences_race.rds")
+
 political_preferences_importances <-
@@
 ```
+
+## Cultural Model
 
@@
         'age_group30-44' = 'Age (30-44)',
-        'age_group45-65' = 'Age (45-65)',
+        'age_group45-65' = 'Age (45-64)',
         'age_group65+' = 'Age (65+)'
@@
 ```
+
+## Political Importances Model
 
@@
 
-@tbl-modelsummarycultural
+## Cultural Model
+
+  @tbl-modelsummarycultural first suggests a that males were less likely to vote for Biden, and more likely to vote for Trump.
+Looking at @fig-agevsgender-1 this is not an obvious correlation to make with the votes for Biden and Trump being close for each age group, with Trump only winning more votes with the male and above 65 demographic.
+However, looking at the 90% confidence intervals in @fig-conf_cultural for the coefficients of that model, specifically that for being male, we can see that the model is 90% confident that this relationship is around strictly negative against voting for Biden.
+We also see the disagreement within the 45+ community that @fig-agevsgender-1 shows, and the agreement within the younger community (18 - 45), with @tbl-modelsummarycultural showing a large correlation between being 30-44 to supporting Biden, with being aged 45-65 or 65+ having very little correlation to support Biden, supported by the 90% confidence interval shown in @fig-conf_cultural for those coefficients spanning both negative and positive, suggesting an uncertain preference.
+Although the model included being: Female, Asian, 18-29, and Agnostic, in the intercept, for @fig-agevsgender-1, @fig-racevsreligion2, and @fig-racevsreligionlarge and @fig-racevsreligion2 respectively, it reasonable to say that each graph suggests that being Female, Asian, 18-29, and Agnostic has a correlation with voting for Biden, each demographic having a significant amount more votes for Biden than Trump.
+
+  Now looking at the coefficients corresponding to race and religion, @tbl-modelsummarycultural suggests that respondents identifying themselves as being Black or Atheist were more likely to have voted for Biden.
+Additionally, as mentioned before those who are Agnostic or Asian as per @tbl-modelsummaryrace, who were included as the intercept are suggested to have voted for Biden.
+The respondents identifying as White, Hispanic, Middle Eastern, Multiracial, and Native American are suggested to be less likely to vote for Biden.
+
+  Overall, this model reinforces our expectations that suggest that the cultural background of a person has a considerable effect on their expected vote.
+Religion and Race especially have considerable leanings to one candidate over the other, suggesting a strong impact of race and religion in political ideals.
+
+## Political Importances Model
+
+  An strong finding that @tbl-modelsummaryimportances displays is household gun ownership's effect on their vote.
+As mentioned in the [model](#Model)'s justification, Trump held a pro-gun-ownership stance, whilst Biden pushed for reform to gun laws.
+@tbl-modelsummaryimportances mirrors these stances, predicting that respondents who own or have people in their household own guns are much more likely to vote for Trump, who aligns with their own personal gun ownership stance.
+This is strongly supported by @fig-confimportances, which shows that the model is 90% confident that the likelihood of a respondent whose household owns guns is between around -0.9 to -1.8, a strongly negative correlation.
+This is supported by @fig-gunownvseducation showing that that most respondents who do not own guns, had voted for Biden, and Trump was only able to secure a majority vote for any level of education, when the respondent's household owned guns.
+
+  Similarly, @tbl-modelsummaryimportances shows a correlation between having student loans and supporting Biden.
+Also mentioned in the [model](#Model)'s justification, the student loan forgiveness stance that more widely benefited current student loan holders was held by Biden.
+This fact is supported by the model that suggests a respondent who has student loans is more likely to support Biden over Trump.
+Being such an significant difference in everyday finances, whether they have to think about paying off student loans or not, it is reasonable to expect the outcome that respondents would want to vote for the candidate who can help take that burden off.
+
+  It is also important to note education's impact on respondent's voting habits.
+The 90% confidence interval the model has for a respondent with no high school education in @fig-confimportances is significantly large, spanning from around -0.6 to +1.6, signifying an uncertainty among that demographic about who to vote for, those with higher education agree more on who to vote for.
+High school graduates slightly favour Biden, and those with University, College, or Post Graduate education are suggested by @tbl-modelsummaryimportances to be more likely to support Biden.
+As this relates to gun ownership as in @fig-gunownvseducation, we can also see here that the impact of owning a gun or not is offset by the respondents education.
+The votes are much closer between Biden and Trump for those with higher education and own guns compared to those without higher education and own guns.
 
 # Discussion
+
+what is done and what we learned:::
+
+In this paper we aimed to verify our expectations that ones cultural background and what one holds to personal importance affect their decision making process when it comes to elections.
+As expected, being of different cultural background impacted who they voted for.
+Exactly what those ideals that they chose to vote with we did not define here, but we know that correlation exists.
+Better defined here is the specific ideals that a respondent holds, modeling gun ownership and student loans and education levels in this paper.
+Also as expected we find that knowing a respondent's opinions on key political topics, in this case gun laws and student loan forgiveness, can allow us to predict who they are voting for, specifically here Trump who was in support of gun ownership laws and Biden who was in support of student loan forgiveness.
+
+weakenesses of the analysis::: TODO also talk about small demographic sizes
 
@@
 
+future steps to learn in the future:::
+
+TODO talk about the change or no change that occurs in the 2024 election, and if respondents of different demographics had changed their minds after Biden's term.
+If people who had their student loans forgiven will still vote for Biden, or vote for Trump after their needs from Biden have already been met.
+
 # Appendix {.unnumbered}
+
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: tbl-modelsummaryrace
+#| tbl-cap: "Explaining whether a voter supported Biden or Trump in 2020 given their race"
+
+# Rename coefficients
+cm <- c(
+        'raceOther' = 'Other Race',
+        'raceWhite' = 'White',
+        'raceNative American' = 'Native American', 
+        'raceMultiracial' = 'Multiracial',
+        'raceMiddle Eastern' = 'Middle Eastern',
+        'raceHispanic' = 'Hispanic',
+        'raceBlack' = 'Black',
+        # Since "Asian" comes first alphabetically, it was factored as '0', being therefore included in the intercept. 
+        # as for this model only includes 
+        '(Intercept)' = 'Asian'
+        )
+
+modelsummary(
+  list(
+    "Support Biden" = political_preferences_race
+  ),
+  statistic = "mad",
+  coef_rename = cm
+  )
+```
 

### 2024-03-17 output/paper.qmd
fix figure crossref and justification formatting, fix grammar mistakes

@@
 
-The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump [cite2020Election].
+The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump \[cite2020Election\].
 Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors.
 We will be using a representative sample of 60000 Americans to discern patterns in these voters' responses and how their personal backgrounds may have influenced their voting decision.
-We will be examining the impact of various factors such as: gender, race, educational background, religon, and personal stances on key political issues such as gun control and student loans.
+We will be examining the impact of various factors such as: gender, race, educational background, religion, and personal stances on key political issues such as gun control and student loans.
 Our main focus is to compare the correlations in the 2022 CES survey data to the 2020 CES survey data, and to see if our estimand, the effect of a voter's cultural background and political stances on their voting decision, has influenced their voting decision in the 2020 election.
@@
 
-The variables we used in our analysis were: - `presvote20post`, renamed to `voted_for` - This data was represented in the form of a numerical variable corresponding to the presidential candidate that voted for in the 2020 eleciton.
+The variables we used in our analysis were: - `presvote20post`, renamed to `voted_for` - This data was represented in the form of a numerical variable corresponding to the presidential candidate that voted for in the 2020 election.
 This was cleaned to limit the options to either "Biden" or "Trump", the two leading candidates in the 2020 and 2024 election.
@@
 This could be attributed to certain parties aligning more to certain religious beliefs.
-To see a bigger insight into this, take a look at @fig-racevsreligionlarge and @fig-racevsreligion2 (figures separated according to respondent count), and notice examples like, 'White Protestants' appear significantly more supportive towards the Republicans as opposed to 'White Catholics' and 'Black Protestants' who appear to be significantly more supportive towards the Democrats.
+To see a bigger insight into this, take a look at @fig-racevsreligionlarge-1, @fig-racevsreligionlarge-2 and @fig-racevsreligion2 (figures separated according to respondent count), and notice examples like, 'White Protestants' appear significantly more supportive towards the Republicans as opposed to 'White Catholics' and 'Black Protestants' who appear to be significantly more supportive towards the Democrats.
 
@@
 
-# Model
+# Model {#model}
 
@@
 
-## Political Importances Model
+## Political Importance Model
 
@@
 We also see the disagreement within the 45+ community that @fig-agevsgender-1 shows, and the agreement within the younger community (18 - 45), with @tbl-modelsummarycultural showing a large correlation between being 30-44 to supporting Biden, with being aged 45-65 or 65+ having very little correlation to support Biden, supported by the 90% confidence interval shown in @fig-conf_cultural for those coefficients spanning both negative and positive, suggesting an uncertain preference.
-Although the model included being: Female, Asian, 18-29, and Agnostic, in the intercept, for @fig-agevsgender-1, @fig-racevsreligion2, and @fig-racevsreligionlarge and @fig-racevsreligion2 respectively, it reasonable to say that each graph suggests that being Female, Asian, 18-29, and Agnostic has a correlation with voting for Biden, each demographic having a significant amount more votes for Biden than Trump.
+Although the model included being: Female, Asian, 18-29, and Agnostic, in the intercept, for @fig-agevsgender-1, @fig-racevsreligion2, and @fig-racevsreligionlarge-1, @fig-racevsreligionlarge-2 and @fig-racevsreligion2 respectively, it reasonable to say that each graph suggests that being Female, Asian, 18-29, and Agnostic has a correlation with voting for Biden, each demographic having a significant amount more votes for Biden than Trump.
 
@@
 
-weakenesses of the analysis::: TODO also talk about small demographic sizes
+weaknesses of the analysis::: TODO also talk about small demographic sizes
 

### 2024-03-17 scripts/04_test_data.R
Add test script

@@ -1,0 +1,18 @@
+#### Preamble ####
+# Purpose: Download US CES
+# Author: Timothius Prajogi, Janssen Myer Rambaud
+# Date: 23 January 2024
+# Contact: tim.prajogi@mail.utoronto.ca
+# License: MIT
+# Prerequisites: 01_download_data, 02_clean_data
+
+#### Workplace setup ####
+
+library(testthat)
+library(tidyverse)
+
+#### Test that the data is as expected ####
+
+# Load in data
+ces2022_1 <- read_csv("output/data/cleaned_ces2022_1.csv")
+ces2022_2 <- read_csv("output/data/cleaned_ces2022_2.csv")

### 2024-03-17 scripts/00_simulate_data.R
Update dates in preamble

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates data for predicting US political support based on demographic and socio-economic factors
 # Author: Janssen Myer Rambaud & Timothius Prajogi
-# Data: March 11 2024
+# Data: March 16 2024
 # Contact: janssen.rambaud@mail.utoronto.ca, timothius.prajogi@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None

### 2024-03-17 scripts/01_download_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Download US CES
 # Author: Timothius Prajogi, Janssen Myer Rambaud
-# Date: 23 January 2024
+# Date: 16 March 2024
 # Contact: tim.prajogi@mail.utoronto.ca
 # License: MIT
 # Prerequisites: none

### 2024-03-17 scripts/02_clean_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Download US CES
+# Purpose: Clean the raw CES data
 # Author: Timothius Prajogi, Janssen Myer Rambaud
-# Date: 23 January 2024
+# Date: 16 March 2024
 # Contact: tim.prajogi@mail.utoronto.ca
 # License: MIT
 # Prerequisites: 01_download_data

### 2024-03-17 scripts/03_model.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Download US CES
+# Purpose: Create GLM on CES cleaned data
 # Author: Timothius Prajogi, Janssen Myer Rambaud
-# Date: 23 January 2024
+# Date: 16 March 2024
 # Contact: tim.prajogi@mail.utoronto.ca
 # License: MIT
 # Prerequisites: 01_download_data, 02_clean_data

### 2024-03-17 scripts/04_test_data.R
add column and some data tests

@@ -16,3 +16,27 @@ library(tidyverse)
 # Load in data
 ces2022_1 <- read_csv("output/data/cleaned_ces2022_1.csv")
 ces2022_2 <- read_csv("output/data/cleaned_ces2022_2.csv")
+
+
+# Test that column types are as expected
+expect_equal(class(ces2022_1$voted_for), "character")
+expect_equal(class(ces2022_1$age_group), "character")
+expect_equal(class(ces2022_1$gender), "character")
+expect_equal(class(ces2022_1$education), "character")
+expect_equal(class(ces2022_1$race), "character")
+expect_equal(class(ces2022_1$religion), "character")
+
+expect_equal(class(ces2022_2$voted_for), "character")
+expect_equal(class(ces2022_2$age_group), "character")
+expect_equal(class(ces2022_2$gender), "character")
+expect_equal(class(ces2022_2$education), "character")
+expect_equal(class(ces2022_2$race), "character")
+expect_equal(class(ces2022_2$religion), "character")
+expect_equal(class(ces2022_2$household_gun_ownership), "character")
+expect_equal(class(ces2022_2$student_loan_status), "character")
+
+
+# Test that data only contains expected values
+expect_equivalent(sort(unique(ces2022_1$voted_for)), sort(c("Biden", "Trump")))
+expect_equivalent(sort(unique(ces2022_1$age_group)), sort(c("18-29", "30-44", "45-64", "65+")))
+

### 2024-03-17 output/paper.qmd

@@
 
-# Discussion
-
-what is done and what we learned:::
+# Discussion {#discussion}
 
@@
 
-weaknesses of the analysis::: TODO also talk about small demographic sizes
+Although there seems to be a strong correlation between if a voter identifies as an other gender to supporting Biden, and a strong correlation between if a voter identifies as an other race to supporting Biden, but these must be considered carefully as within the sample (with seed 302), only 5 and 1 voters identified in those categories respectively.
+There were various other demographics in the dataset that had relatively small sample sizes, such as: `race` where non black, white, or hispanic respondents made up a small portion of the sample, or non Catholic, Protestant, or Atheist respondents in `religion`.
+To get a more accurate reading in regards to these features, it would be necessary to train a model which includes a larger sample size of these demographics or to find a way to account for these values while still making use of them.
+This was a scenario that one may expect to see as there can always be underrepresented demographics in a 60000 respondent sized sample.
+This is a big limitation to our models, and is something that should be considered when interpreting the results.
 
-Although there seems to be a strong correlation between if a voter identifies as an other gender to supporting Biden, and a strong correlation between if a voter identifies as an other race to supporting Biden, but these must be considered carefully as within the sample (with seed 302), only 5 and 1 voters identified in those categories respectively.
-To get a more accurate reading in regards to these features, it would be necessary to train a model which includes much more of these demographics.
-
-future steps to learn in the future:::
-
-TODO talk about the change or no change that occurs in the 2024 election, and if respondents of different demographics had changed their minds after Biden's term.
-If people who had their student loans forgiven will still vote for Biden, or vote for Trump after their needs from Biden have already been met.
+For future steps, we hope to be able to observe any changes that arise in the upcoming 2024 election.
+Since the results we have interpreted from the 2022 CES dataset are based on voter's stances at the time, there could be a possibility that the voter's stances have changed nearing the end of Biden's term.
+It would also be good to see if the results from the our models are consistent with the results from the upcoming 2024 election.
+Should there be any consistency, would it raise the question of whether the voter's stances are more influenced by their cultural background and personal stances, or if their support is more influenced by the continuously changing platform that each party stands on.
+There are some areas that would be interesting to observe if Harvard were to conduct a 2024 CES survey, such as if previous Biden voters would still vote for him considering how he has handled the student loan forgiveness situation.
+Would a voter who has their student loans forgiven this term still vote for Biden since their needs have already been met?
+It would also be note worthy to take a look at how stances have changed on household gun ownership and if it has affected their support in the presidential election.
+As more shootings in the country are recorded, will the number of gun owners increase?
+Will this affect which party they support, considering the different stances on gun ownership laws?
+These are all questions that we hope could be answered after the outcome of the 2024 election.
 

### 2024-03-17 output/paper.qmd
Properly cite

@@
 
-The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump \[cite2020Election\].
+The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump @cite2020Election.
 Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors.
@@
 Although there seems to be a strong correlation between if a voter identifies as an other gender to supporting Biden, and a strong correlation between if a voter identifies as an other race to supporting Biden, but these must be considered carefully as within the sample (with seed 302), only 5 and 1 voters identified in those categories respectively.
-There were various other demographics in the dataset that had relatively small sample sizes, such as: `race` where non black, white, or hispanic respondents made up a small portion of the sample, or non Catholic, Protestant, or Atheist respondents in `religion`.
+There were various other demographics in the dataset that had relatively small sample sizes, such as: `race` where non Black, White, or Hispanic respondents made up a small portion of the sample, or non Catholic, Protestant, or Atheist respondents in `religion`.
 To get a more accurate reading in regards to these features, it would be necessary to train a model which includes a larger sample size of these demographics or to find a way to account for these values while still making use of them.

### 2024-03-17 output/paper.qmd
Try to update formatting

@@
 
-Although there seems to be a strong correlation between if a voter identifies as an other gender to supporting Biden, and a strong correlation between if a voter identifies as an other race to supporting Biden, but these must be considered carefully as within the sample (with seed 302), only 5 and 1 voters identified in those categories respectively.
+Although there seems to be a strong correlation between if a voter identifies as an other gender to supporting Biden, and a strong correlation between if a voter identifies as an other race to supporting Biden as suggested by the 90% confidence interval for each shown in @fig-confother, but these must be considered carefully as within the sample (with seed 302), only 5 and 1 voters identified in those categories respectively.
 There were various other demographics in the dataset that had relatively small sample sizes, such as: `race` where non Black, White, or Hispanic respondents made up a small portion of the sample, or non Catholic, Protestant, or Atheist respondents in `religion`.
@@
 
+\newpage
 # Appendix {.unnumbered}

### 2024-03-17 output/paper.qmd
updated referencing to arrow and fixed citation

@@
 
-The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump @cite2020Election.
+The topic of concern is the 2020 United States presidential election held on November 3rd, 2020, where we saw Democratic Party leader Joe Biden win against incumbent president Donald Trump [@cite2020Election].
 Using the 2022 Cooperative Election Study (CES) survey data, we aim to predict who a voter supported in the 2020 election based on their cultural background, political stances, and other factors.
@@
 
-For additional assitance with report generation the `knitr` [@citeknitr] package was used.
+For additional assistance with data storage, the `arrow`[@citearrow] was used and for report generation, the `knitr` [@citeknitr] package was used.
 

### 2024-03-17 scripts/02_clean_data.R
added parquet

@@ -9,6 +9,7 @@
 #### Workplace setup ####
 
 library(tidyverse)
+library(arrow)
 
 #### Clean CES Data ####
 
@@ -134,6 +135,7 @@ cleaned_ces2022_1 <-
   select(voted_for, age_group, gender, education, race, religion)
 
 write_csv(cleaned_ces2022_1, "output/data/cleaned_ces2022_1.csv")
+write_parquet(cleaned_ces2022_1, "output/data/cleaned_ces2022_1.parquet")
 
 # Gun ownership and education loan status have around 10,000 missing values,
 # so we will first store the data with columns not including gunown and edloan, 
@@ -167,5 +169,6 @@ cleaned_ces2022_2 <-
   select(-gunown, -edloan)
 
 write_csv(cleaned_ces2022_2, "output/data/cleaned_ces2022_2.csv")
+write_parquet(cleaned_ces2022_2, "output/data/cleaned_ces2022_2.parquet")
   
-  
+

### 2024-03-17 output/paper.qmd
added parquet
Final changes

--- 2024-03-17 output/paper.qmd
@@
 Looking at @fig-agevsgender-1 this is not an obvious correlation to make with the votes for Biden and Trump being close for each age group, with Trump only winning more votes with the male and above 65 demographic.
-However, looking at the 90% confidence intervals in @fig-conf_cultural for the coefficients of that model, specifically that for being male, we can see that the model is 90% confident that this relationship is around strictly negative against voting for Biden.
-We also see the disagreement within the 45+ community that @fig-agevsgender-1 shows, and the agreement within the younger community (18 - 45), with @tbl-modelsummarycultural showing a large correlation between being 30-44 to supporting Biden, with being aged 45-65 or 65+ having very little correlation to support Biden, supported by the 90% confidence interval shown in @fig-conf_cultural for those coefficients spanning both negative and positive, suggesting an uncertain preference.
+However, looking at the 90% confidence intervals in @fig-conf_cultural for the coefficients of that model, specifically that for being male, we can see that the model is 90% confident that this relationship is strictly negative, signifying they are against voting for Biden.
+The model also suggests disagreement within the 45+ community and the agreement within the younger community (18 - 44), something that is evident in @fig-agevsgender-1.
+@tbl-modelsummarycultural also shows a large correlation between being 30-44 to supporting Biden, with being aged 45-65 or 65+ having very little correlation to support Biden, supported by the 90% confidence interval shown in @fig-conf_cultural for those coefficients spanning both negative and positive, suggesting an uncertain preference.
 Although the model included being: Female, Asian, 18-29, and Agnostic, in the intercept, for @fig-agevsgender-1, @fig-racevsreligion2, and @fig-racevsreligionlarge-1, @fig-racevsreligionlarge-2 and @fig-racevsreligion2 respectively, it reasonable to say that each graph suggests that being Female, Asian, 18-29, and Agnostic has a correlation with voting for Biden, each demographic having a significant amount more votes for Biden than Trump.

### 2024-03-17 output/paper.qmd

@@
 
-For additional assitance with report generation the `knitr` [@citeknitr] package was used.
+For additional assistance with data storage, the `arrow` [@citearrow] was used and for report generation, the `knitr` [@citeknitr] package was used.
 

### 2024-03-17 output/paper.qmd
Final changes

@@
 However, looking at the 90% confidence intervals in @fig-conf_cultural for the coefficients of that model, specifically that for being male, we can see that the model is 90% confident that this relationship is strictly negative, signifying they are against voting for Biden.
-The model also suggests disagreement within the 45+ community and the agreement within the younger community (18 - 44), something that is evident in @fig-agevsgender-1.
-@tbl-modelsummarycultural also shows a large correlation between being 30-44 to supporting Biden, with being aged 45-65 or 65+ having very little correlation to support Biden, supported by the 90% confidence interval shown in @fig-conf_cultural for those coefficients spanning both negative and positive, suggesting an uncertain preference.
+The model suggests disagreement within the 45+ community and the agreement within the younger community (18 - 44), something that is evident in @fig-agevsgender-1.
+From @tbl-modelsummarycultural we also see a large correlation between being 30-44 to supporting Biden, with being aged 45-65 or 65+ having very little correlation in supporting Biden, supported by the 90% confidence interval shown in @fig-conf_cultural for those coefficients spanning both negative and positive, suggesting an uncertain preference.
 Although the model included being: Female, Asian, 18-29, and Agnostic, in the intercept, for @fig-agevsgender-1, @fig-racevsreligion2, and @fig-racevsreligionlarge-1, @fig-racevsreligionlarge-2 and @fig-racevsreligion2 respectively, it reasonable to say that each graph suggests that being Female, Asian, 18-29, and Agnostic has a correlation with voting for Biden, each demographic having a significant amount more votes for Biden than Trump.
